\documentclass[a4paper, 12pt]{article}
\input{../preamble.tex}
\title{MATH 26200: Point-Set Topology \\ \large Take-home Midterm Exam}
\date{31 Jan 2024}
\author{Hung Le Tran}
\begin{document}
\maketitle
\setcounter{section}{4}
\textbf{Textbook:} Munkres, \textit{Topology}  
\begin{problem} [\redtext{done}]
    Let $\bbr$ denote the real numbers. Give $\bbr$ the topology in which the closed sets (other than all of $\bbr$) are the finite subsets. Verify that this is a topology, and prove that if $p(x)$ is a polynomial (with real coefficients), the function $x \mapsto p(x)$ is continuous in this topology.
\end{problem}
\begin{solution}
    We verify the properties of a topology through its closed sets:
    \begin{enumerate} [1.]
        \item $\bbr$ is closed by hypothesis. $\emptyset$ has $0 < \infty$ elements, so is closed.
        \item Arbitrary intersections of closed $\{K_\alpha\}_{\alpha \in A}$, $\bigcap K_\alpha$, has at most $\abs{K_\alpha}$ elements for some $\alpha \in A$, but $\abs{K_\alpha} < \infty$ so $\bigcap K_\alpha$ is also closed.
        \item Finite unions of clsoed $\{K_i\}_{i \in [N]}$ has at most $\sum_{i=1}^{N} \abs{K_i}  < \infty$, so is also closed.
    \end{enumerate}

    It follows that this is indeed a topology.

    Now, let $p(x)$ be any polynomial in $x$. Let $N = \deg (p(x)) \in \bbz_{\geq 0}$. To show that $p$ is continuous, want to show preimages of closed sets are closed. Let $K = \{y_1, \ldots, y_n\}$ be any closed set. Then:
    \begin{align*}
    p^{-1}(K) &= \bigcup_{i \in [n]} p^{-1}(y_i) \\
    &= \bigcup_{i \in [n]} \{x : p(x) - y_i = 0\}
    \end{align*}
    And we have $\deg (p(x)) = N \implies \deg(p(x) - y_i) = N $ so it has at most $N$ roots. Therefore: 
    \begin{equation*}
    \abs{p^{-1}(K)} \leq Nn < \infty    
    \end{equation*}
    so $p^{-1}(K)$ is closed, as required.
\end{solution}

\begin{problem} [18.1 \redtext{done}]
    Prove that for functions $f: \bbr \to \bbr$, the $\epsilon - \delta$ definition of continuity implies the open set definition.
\end{problem}
\begin{solution}
    WTS if a function $f: \bbr \to \bbr$ satisfies the $\epsilon - \delta$ condition, then it is continuous in the open set definition.

    Take any $W \subset \bbr$ open, want to prove that $f^{-1}(W)$ is open. Take $x \in f^{-1}(W)$, i.e., $f(x) \in W$. $f(x) \in W$ open, so there exists a basis $f(x) \in B(f(x'), \epsilon') \subset W$. Define $\epsilon \coloneqq \frac{1}{2} \min\{\epsilon' - \abs{f(x) - f(x')}, \abs{f(x) - f(x')}\} > 0$, then $f(x) \in B(f(x), \epsilon) \subset B(f(x'), \epsilon') \subset W$.
    
    From the $\epsilon - \delta$ condition, it follows that there exists some $\delta > 0$ such that $f(B(x, \delta)) \subset B(f(x), \epsilon) \subset W$. It therefore follows that $B(x, \delta) \subset f^{-1}(W)$. We can do this for all $x \in f^{-1}(W)$, so it is open. $f$ is therefore continuous in the open set definition.
\end{solution}

\begin{problem} [19.8 \redtext{done}]
    Given sequences $(a_1, a_2, \ldots)$ and $(b_1, b_2, \ldots)$ of real numbers with $a_i > 0$ for all $i$, define $h: \bbr^\omega \to \bbr^\omega$ by the equaiton:\begin{equation*}
    h((x_1, x_2, \ldots)) = (a_1 x_1 + b_1, a_2x_2 + b_2, \ldots)
    \end{equation*}

Show that if $\bbr^\omega$ is given the product topology, $h$ is a homeomorphism of $\bbr^\omega$ with itself. What happens if $\bbr^\omega$ is given the box topology?
\end{problem}
\begin{solution}
    To show that $h$ is homeo, we want to show that it is bijective, continuous and that $h^{-1}$ is also continuous.

    \begin{enumerate} [1.]
    \item We write \begin{equation*}
        g((y_1, y_2, \ldots)) = \left(\frac{y_1 - b_1}{a_1}, \frac{y_2 - b_2}{a_2}, \ldots\right)  
    \end{equation*}
    This map is well-defined since $a_i > 0 \forall i \in \bbn$. It's trivial that $h \circ g = g \circ h = id$. It follows that $h$ is a bijection, with inverse $h^{-1} = g$.
    \item WTS each of $h$'s coordinate functions is continuous. Define, for each $i$, $f_i: \bbr \to \bbr, f_i(x) = a_i x + b_i$ then $f_i$ is trivially continuous. Then $h_i(x) = f_i \circ \pi_i$ is a composition of 2 continuous functions, and is therefore also continuous. Since each of $h$'s coordinate functions is continuous, $h$ is also continuous.
    \item $g = h^{-1}$ is continuous because of the same reason, each $g_i$ is similarly continuous.
    \end{enumerate}
    Therefore indeed $h$ is homeo.

    In the box topology, claim that $h$ is also a homeo. First, trivially, it is still a bijection. 
    
    Take the typical basis element $\prod_{i \in \bbn} U_i$, then we have 
    \begin{equation*}
        h^{-1}(\prod_{i \in \bbn} U_i) = \prod_{i \in \bbn} h_i^{-1}\left(U_i\right) 
    \end{equation*}
    Each $h_i$, as mentioned above, is continuous so $h_i^{-1}(U_i)$ is open in $\bbr$. Therefore $\prod_{i \in \bbn} h_i^{-1}\left(U_i\right)$ is open in the box topology, which implies $h$ is continuous.

    The proof for $g$ is continuous is similar, since each $g_i$ is similarly continuous. It follows that $h$ is a homeo in the box topology as well.
\end{solution}

\begin{problem} [23.5 \redtext{done}]
    A space is \textbf{totally disconnected} if its only connected subspaces are one-point sets. Show that if $X$ has the discrete topology, then $X$ is totally disconnected. Does the converse hold?
\end{problem}
\begin{solution}
    By hypothesis, $X$ has the discrete topology. One-point sets are clearly connected, since there can't be a separation with 2 non-empty disjoint clopen subsets, which would make the number of elements in the set $\geq 2$. Suppose $K$ with $\abs{K} \geq 2$ is also connected. Then there exists $a \in K$, and $K = \{a\} \sqcup (K - \{a\})$ is a separation of $K$ into 2 non-empty closed sets, so $K$ is not connected, \contra.

    It follows that the only connected subspaces are one-point sets (technically $\emptyset$ is also connected, but I figure the connected subspaces should be non-trivial), so $X$ is totally disconnected as required.

    % WTS the converse holds. Note that one-point sets are always connected, since its number of elements does not allow it to have a separation with 2 non-empty disjoint subsets.

    % If $X$ has 1 point then it holds trivially that the only (non-trivial) connected subspaces are one-point sets — there are no other sets.

    % If $X$ has more than 2 points, then for all $x \in X$, choose $y \in x, y \in X$, the set $\{x , y\}$ is not connected, meaning there exists some separation with 2 non-empty, disjoint, clopen subsets. The subsets can therefore only be $\{x\}$ and $\{y\}$, and they are clopen. Therefore $\{x\}$ is open for all $x \in X$, so $X$ has the discrete topology.

    The converse does not hold. Note that one-point sets are always connected, since its number of elements does not allow it to have a separation with 2 non-empty disjoint subsets. 
    
    Take $\bbq$ in the subspace topology, $\bbq \subset \bbr$. This is not the discrete topology, for each basis element $\bbq \cap B(x, r)$ contains infinitely many points. However, we claim that $\bbq$ is indeed totally disconnected. Suppose for sake of contradiction that there exists $U \subset \bbq$ connected such that $\abs{U} \geq 2$. Then we can find $q_1, q_2 \in U; q_1 < q_2$. We can then exhibit a separation of $U$:
    \begin{equation*}
    U = (U \cap (-\infty, p) \sqcup (U \cap (p, +\infty))
    \end{equation*}
    where $p$ is some irrational such that $q_1 < p < q_2$. They are clearly disjoint and open, and non-empty since $q_1$ and $q_2$ are respectively in them. So $U$ is not connected. $\bbq$ is therefore totally disconnected, but doesn't have the discrete topology.
\end{solution}

\begin{problem} [24.1 \redtext{done}]
    \begin{enumerate}
    \item[] 
    \item Show that no two of the spaces $(0, 1), (0, 1]$ and $[0, 1]$ are homeomorphic. [Hint: What happens if you remove a point from each of these spaces?]
    \item Suppose that there exist embeddings $f: X \to Y$ and $g: Y \to X$. Show by means of an example that $X$ and $Y$ need not be homeomorphic.
    \item Show $\bbr^n$ and $\bbr$ are not homeomorphic if $n > 1$.
\end{enumerate}
\end{problem}
\begin{solution}
\textbf{(a)} Removing any point $\alpha$ from $(0, 1)$ makes it a disconnected space: $(0, 1) = (0, \alpha) \sqcup (\alpha, 1)$, while removing $1$ from $(0, 1]$ keeps the space connected ($(0, 1)$ is connected) and removing $1$ from $[0, 1]$ also keeps the space connected ($[0, 1)$ is connected), so $(0, 1)$ is not homeomorphic to $(0, 1]$ nor is it homeomorphic to $[0, 1]$.

Removing any 2 distinct points from $(0, 1]$ makes it a disconnected space (removing $1$ and another point $\alpha \in (0, 1)$ makes it disconnected: $(0, \alpha) \sqcup (\alpha, 1)$; and removing $\alpha, \beta \in (0, 1)$ yields $(0, \alpha) \sqcup (\alpha, 1]$), while removing $0$ and $1$ from $[0, 1]$ keeps the space connected ($(0, 1)$ is connected), so $(0, 1]$ is not homeomorphic to $[0, 1]$.

Why this ``removing'' reasoning work is that suppose there exists a homeomorphism $f: (0, 1] \to (0, 1)$. Then $f((0, 1)) = (0, 1) - \{f(1)\}$; the LHS is connected (since $f$ is continuous and $(0, 1)$ is connected) while the RHS is not, \contra.

Similar for the removing-2-point case.

\textbf{(b)} Take $X = (0, 1), Y = [0, 1]$, then $f: X \to Y, f(x) = \frac{x}{2}$ and $g: Y \to X, g(y) = \frac{y+1}{4}$ are embeddings, but $(0, 1)$ and $[0, 1]$ are not homeomorphic as abovementioned.

\textbf{(c)} $n > 1$. 

Removing any point $\alpha$ from $\bbr$ makes it a disconnected space: $(- \infty, \alpha) \sqcup (\alpha, \infty)$.

However, removing $0 \in \bbr^n$ keeps it connected. So $\bbr^n - \{0\}$ is not homeomorphic to $\bbr$.
\end{solution}

\begin{problem} [24.10 \redtext{done}]
    Show that if $U$ is an open connected subspace of $\R^2$ then $U$ is path connected. [Hint: Show that given $x_0 \in U$, the set of points that can be joined to $x_0$ by a path in $U$ is both open and closed in $U$.]
\end{problem}
\begin{solution}
    Take $x_0 \in U$. Let $A = \{x \in U: \:\text{there exists a path}\: x_0 \to x\}$. Then $x_0 \in A$, so $A$ is non-empty. WTS $A$ is clopen in $U$.

    Take any $x \in A$, then since $U$ is open, there exists some $B(x, \epsilon)$ (can always center the ball) such that $x \in B(x, \epsilon) \subset U$. Take any $y \in B(x, \epsilon)$, then since the ball is convex, the segment $x \to y$ is contained in $B(x, \epsilon)$, and therefore in $U$. Therefore $f(t) = x + t(y-x)$ is a path from $x \to y$. $x \in A$, so there exists a path from $x_0 \to x$. Concatenate these 2 paths and we get a path from $x_0 \to y$, so $y \in A$. This works for any $y \in B(x, \epsilon)$, so $B(x, \epsilon) \subset A$. This works for any $x \in A$, so $A$ is open.

    $A$ is also closed, take $X - A = \{x \in U: \:\text{there is no path}\: x_0 \to x\}$. Take any $x \in X - A$, then since $U$ is open, there exists some $B(x, \epsilon)$ (can always center the ball) such that $x \in B(x, \epsilon) \subset U$. Take any $y \in B(x, \epsilon)$, then similarly, $f(t) = x + t(y-x)$ is a path from $x \to y$. We can conclude that $y \in X -A$, because if $y \in A$, meaning there's a path from $x_0 \to y$, then one can concatenate that path with the path from $x \to y$ in reverse to get a path from $x_0 \to x$, but $x \not \in A$, so it would be a contradiction. Therefore for any $y \in B(x, \epsilon)$, $y \in X - A \implies B(x, \epsilon) \subset X - A$. This works for any $x \in X - A$, so $X - A$ is open, so $A$ is closed.

    $A$ is clopen and non-empty, $A \subset U$ connected, so $A = U$. There is a path from $x_0$ to all $x \in U$. Then for all $x, y \in U$, concatenate path $x_0\to x$ in reverse to path $x_0\to y$, and we get path $x \to y$. So $U$ is path-connected.
\end{solution}

\begin{problem} [26.11 \redtext{done}]
Let $X$ be a compact Hausdorff space. Let $\cala$ be a collection of closed connected subsets of $X$ that is simply ordered by proper inclusion. Then \begin{equation*}
Y = \bigcap_{A \in \cala} A
\end{equation*}
is connected. [Hint: if $C \cup D$ is a separation of $Y$, choose disjoint open sets $U$ and $V$ of $X$ containing $C$ and $D$, respectively, and show that $\bigcap_{A \in \cala}(A - (U \cup V))$ is not empty.]
\end{problem}
\begin{solution}
    Each $A$ is closed so $Y = \bigcap_{A \in \cala} A$ is also closed in $X$.

    Suppose, for the sake of contradiction, that $Y$ is not connected, i.e., that there exists separation $Y = C \sqcup D$ of $C, D$ non-empty, disjoint, clopen sets in $Y$. $C, D$ are closed in $Y$, $Y$ closed in $X$ so they are closed in $X$.

    We use a fact, shown in class, that compact Hausdorff spaces are normal. So $X$ is normal. Therefore, since $C, D$ are closed and disjoint in $X$, there exists open $U, V \subset X$  such that $C \subset U, D \subset V$ and $U \cap V = \emptyset$.

    Then, for each $A \in \cala$, claim that $(A \cap U) \cup (A \cap V) \neq A$. Suppose not, that $(A \cap U) \cup (A \cap V) = A$, then $U \cap V = \emptyset$ implies that $(A \cap U) \cap (A \cap V) = \emptyset$, with both $(A \cap U), (A \cap V)$ open in $A$, and $A \cap U \supset C$, $A \cap V \supset D$ so they are both non-empty. Hence we get a separation of $A$, but $A$ is connected, \contra.

    It follows that $(A \cap U) \cup (A \cap V) \neq A \implies A - (A\cap U) \cup (A \cap V) = A - U \cup V \neq \emptyset$ for all $A \in \cala$.

    $U \cup V$ is open in $X$, so $A - U \cup V$ is closed in $X$. $X$ is compact, so $A - U \cup V$ is compact. $\cala$ is a collection of $A$ ordered by proper inclusion, so $\{A - U\cup V\}_{A \in \cala} $ is a collection of non-empty, compact, subsets ordered by proper inclusion.

    It then follows that \begin{equation*}
    \bigcap_{A \in \cala} (A - U\cup V) \neq \emptyset
    \end{equation*}

    However, \begin{equation*}
        \bigcap_{A \in \cala} (A - U\cup V) = \left(\bigcap_{A \in \cala} A\right) - U \cup V = Y - U \cup V = C \cup D - U \cup V = \emptyset, \contra
    \end{equation*}

    It follows that $Y$ is connected.
\end{solution}

\end{document}